---
title: "E-Net Comparison"
author: "S. Eanes"
date: "2/21/2020"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---


```{r import, warning=F, include=F}
library(glmnet)
library(glmnetUtils)
library(plotmo)
library(tidyverse)
library(MLmetrics)
```

## Setup data

Reserve 25% of the data for validation to comparse different modeling techniques.

```{r data}
coral <- read.csv("C:/Users/shado/Documents/School/Fossil Coral/data/coral_3weighted.csv")
test.prop <- .75

#omit data with no response
coral <- coral[!is.na(coral$U238),]

#select three largest species
genus_trim <- c("Acropora", "Porites", "Favia")
#remove coral with age > 10
coral <- coral[which(coral$Genus %in% genus_trim),]  %>% dplyr::filter(Age < 10)

#clean up a nice dataframe
coral.df <- coral %>% mutate(Temperature = Temp) %>%
  select(U238,pH,TAlk,Salinity,Temperature,OmegaA,TCO2) %>%
  scale() %>%
  data.frame()
nrow(coral.df)
```

## Compute Lasso RMSE

```{r lasso}
##LASSO
lasso <- function() {
  lambda.grid <- exp(seq(-7.5,0,length=100))
  cv.lasso <- cv.glmnet(U238 ~ ., data=coral.df,alpha=1,lambda=lambda.grid,intercept=F)
  #plot(cv.lasso)

  mod.lasso1.grid <- glmnet(U238 ~ ., data=coral.df, alpha=1,lambda=lambda.grid,intercept=F)
  #uncomment and run alone for plots
  #plot_glmnet(mod.lasso1.grid,"lambda",label=8)

  mod.lasso.min <- glmnet(U238 ~ ., data=coral.df,alpha=1,lambda=cv.lasso$lambda.min,intercept=F)
  mod.lasso.1se <- glmnet(U238 ~ ., data=coral.df,alpha=1,lambda=cv.lasso$lambda.1se,intercept=F)
  cv.lasso$lambda.1se; cv.lasso$lambda.min

  p1 <- predict(mod.lasso.min,coral.test)
  p2 <- predict(mod.lasso.1se,coral.test)

  lmse.min <- RMSE(p1,coral.test$U238)
  lmse.1se <- RMSE(p2,coral.test$U238) 
  c(lmse.min,lmse.1se)
}
```

## Compute Ridge RMSE

```{r ridge}
#RIDGE
ridge <- function() {
  lambda.grid <- exp(seq(-8,5,length=100))
  cv.ridge <- cv.glmnet(U238 ~ ., data=coral.df,alpha=0,lambda=lambda.grid,intercept=F)
  #plot(cv.ridge)

  mod.ridge1.grid <- glmnet(U238 ~ ., data=coral.df, alpha=0,lambda=lambda.grid,intercept=F)
  #plot_glmnet(mod.ridge1.grid,"lambda",label=8)

  mod.ridge.min <- glmnet(U238 ~ ., data=coral.df,alpha=0,lambda=cv.ridge$lambda.min,intercept=F)
  mod.ridge.1se <- glmnet(U238 ~ ., data=coral.df,alpha=0,lambda=cv.ridge$lambda.1se,intercept=F)
  cv.ridge$lambda.min; cv.ridge$lambda.1se

  p1 <- predict(mod.ridge.min,coral.test)
  p2 <- predict(mod.ridge.1se,coral.test)

  rmse.min <- RMSE(p1,coral.test$U238)
  rmse.1se <- RMSE(p2,coral.test$U238)
  c(rmse.min,rmse.1se)
}
```

## Compute Elastic Net ($\alpha=.5$) RMSE

```{r elastic-net}

#ELASTIC-NET
enet <- function(a1=.5){
  lambda.grid <- exp(seq(-10,2,length=100))
  cv.net <- cv.glmnet(U238 ~ ., data=coral.df, alpha=a1,lambda=lambda.grid,intercept=F)
  #plot(cv.net)

  mod.enet.grid <- glmnet(U238 ~ ., data=coral.df, alpha=a1,lambda=lambda.grid,intercept=F)
  #plot_glmnet(mod.enet.grid,"lambda",label=8)

  mod.enet.min <- glmnet(U238 ~ ., data=coral.df,alpha=a1,lambda=cv.net$lambda.min,intercept=F)
  mod.enet.1se <- glmnet(U238 ~ ., data=coral.df,alpha=a1,lambda=cv.net$lambda.1se,intercept=F)

  p1 <- predict(mod.enet.min,coral.test) 
  p2 <- predict(mod.enet.1se,coral.test)

  emse.min <- RMSE(p1,coral.test$U238)
  emse.1se <- RMSE(p2,coral.test$U238)
  c(emse.min,emse.1se)
}
```

## Compute many times

Cross-validation can produce slightly different results from run to run, and the random split of test-validation can also produce differing results. To adjust for this, we do the test-train split 10 times, and compute the above regularized regression rmse 100 times for each test-train split, averaging the results.

```{r}
wrapper_func <- function(){
  lvals <- lasso()
  rvals <- ridge()
  evals <- enet()
  c(lvals,rvals,evals)
}
```

```{r run}
n <- 100
resamp <- 10
averages <- data.frame(matrix(rep(0,resamp*6),ncol=6))
colnames(averages) <- c("LMSE.min","LMSE.1se","RMSE.min","RMSE.1se","EMSE.min","EMSE.1se")
  
for(i in 1:resamp){
  print(i)
  MSEs <- data.frame(matrix(rep(0,n*6),ncol=6))
  
  samp <- sample(1:10,nrow(coral.df),replace=T)==i
  coral.test <- coral.df[samp,]
  coral.df <- coral.df[!samp,]
  
  for(j in 1:n){
    MSEs[j,] <- wrapper_func()
  }
  averages[i,] <- colMeans(MSEs)
}
```

## Comparison RMSE

```{r}
sort(colMeans(averages))
```

### Find an optimal alpha

```{r cv-alpha}
#CV alpha param

lambda.grid <- exp(seq(-10,2,length=100))
mod.net <- cva.glmnet(U238 ~ ., data=coral.df,lambda=lambda.grid)
ALPHA <- mod.net$alpha
mod.net$modlist[[1]]$cvm
a1 <- ALPHA[which.min(sapply(mod.net$modlist, function(mod) {print(cbind(min(mod$cvm),lambda.grid[which.min(mod$cvm)]));min(mod$cvm)}))]
a1
```