---
title: "Getting Started"
author: "S. Eanes"
date: "12/27/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r import, warning=F, include=F}
library(tidyverse)
```

## Importing the Data

Data is contained in data/coral_3weighted.csv. This contains all the computed columns, but still needs to be filtered down. We remove NA U238 samples, include on samples of genus acropora nad porites, remove samples older than 10k years, and also remove all samples with a reported calcite value greater than 1. This should result in 700 rows in the data. This final data is stored in data/final_sample.csv.

```{r data}
coral <- read.csv("~/School/Fossil Coral/data/coral_3weighted.csv")

#omit data with no response
coral <- coral[!is.na(coral$U238),]

#select three largest species
genus_trim <- c("Acropora", "Porites")
#remove coral with age > 10
coral <- coral[which(coral$Genus %in% genus_trim),]  %>% dplyr::filter(Age < 10)
coral <- coral[(coral$Calcite <= 1 | is.na(coral$Calcite)),]

#clean up a nice dataframe, change U238 to correct units
coral.df <- coral %>% mutate(Temperature = Temp, U238=U238*.421) %>%
  select(U238,pH,TAlk,Salinity,Temperature,OmegaA,TCO2,Genus) %>%
  data.frame

nrow(coral.df)
```

## Basic Modeling

We can quickly reproduce some of the linear models from Patterson et. al.

```{r temp}
coral.por <- coral.df %>% filter(Genus=="Porites")
coral.acr <- coral.df %>% filter(Genus=="Acropora")
nrow(coral.por)+nrow(coral.acr)==nrow(coral.df)

temp.por <- lm(U238 ~ Temperature, data = coral.por)
summary(temp.por)

temp.acr <- lm(U238 ~ Temperature, data = coral.acr)
summary(temp.acr)

temp <- lm(U238 ~ Temperature, data = coral.df)
summary(temp)
```

```{r}
tsal.por <- lm(U238 ~ Temperature + Salinity, data = coral.por)
summary(tsal.por)

tsal.acr <- lm(U238 ~ Temperature + Salinity, data = coral.acr)
summary(tsal.acr)

tsal <- lm(U238 ~ Temperature + Salinity, data = coral.df)
summary(tsal)
```

