---
title: "Residual Analysis"
author: "S. Eanes"
date: "11/25/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r import, warning=F, include=F}
library(glmnet)
library(glmnetUtils)
library(plotmo)
library(tidyverse)
library(MLmetrics)
library(lmtest)
```

## Setup data

```{r data}
coral <- read.csv("~/School/Fossil Coral/data/coral_3weighted.csv")

#omit data with no response
coral <- coral[!is.na(coral$U238),]

#select three largest species
genus_trim <- c("Acropora", "Porites")
#remove coral with age > 10
coral <- coral[which(coral$Genus %in% genus_trim),]  %>% dplyr::filter(Age < 10)
coral <- coral[(coral$Calcite <= 1 | is.na(coral$Calcite)),]

#clean up a nice dataframe
coral.df <- coral %>% mutate(Temperature = Temp, U238=U238*.421) %>%
  select(U238,pH,TAlk,Salinity,Temperature,OmegaA,TCO2) %>%
  data.frame()

nrow(coral.df)
```

```{r}
coral.summary =coral %>% mutate(Temperature = Temp, U238=U238*.421) %>%
  select(U238,Age,Calcite,U234delta,pH,TAlk,Salinity,Temperature,OmegaA,TCO2) %>%
  data.frame()

summary <- coral.summary %>%
  gather %>% group_by(key) %>%
  summarise("5%"=quantile(value,.05,na.rm=T),
            median=median(value,na.rm=T),
            mean=mean(value,na.rm=T),
            "95%"=quantile(value,.95,na.rm=T)) %>%
  as.data.frame()
cbind(c(summary[,1]),round(summary[,-1],2))
  
```

## Data Distribution

Plot the distributions of the predictors. 
```{r distributions}
ggplot(gather(coral.df),aes(value))+
  geom_histogram(bins=10)+
  facet_wrap(~key, scales = 'free_x')
```

## Residaul Analysis

Residuals vs fitted looks perfectly normal - no signs of non-linearity or heteroskedasticity.

As can be seen in the normal q-q plot, there is some non-normality. In the KDE plot below, and with the shapiro test we can confirm that there is some right skew. However, with a large number of samples like this analysis, linear regression is quite robust to violations of the normality assumption. Furthermore, by excluding high percent calcite samples, we are confident that the samples in this analysis are of high quality. 

Scale-location shows that there is perhaps some slight increase in the magnitude of standardized residuals as the fitted value increases but not to an extent that is a clear violation of assumptions. 

There are no conerning observations when looking at cook's distance in our residuals vs leverage plot.

```{r residuals}
options(scipen=999)
final_model <- lm(U238 ~ pH + Temperature + OmegaA + TAlk, data=coral.df)
summary(final_model)
plot(final_model)
```

```{r additional}
#normal Q-Q additional analysis
par(mfrow=c(1,1))
d<-density(final_model[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
shapiro.test(final_model[['residuals']])
```
