---
title: "Residual Analysis"
author: "S. Eanes"
date: "11/25/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r import, warning=F, include=F}
library(glmnet)
library(glmnetUtils)
library(plotmo)
library(tidyverse)
library(MLmetrics)
library(lmtest)
```

## Setup data

```{r data}
coral <- read.csv("C:/Users/shado/Documents/School/Fossil Coral/data/coral_3weighted.csv")

#omit data with no response
coral <- coral[!is.na(coral$U238),]

#select three largest species
genus_trim <- c("Acropora", "Porites")
#remove coral with age > 10
coral <- coral[which(coral$Genus %in% genus_trim),]  %>% dplyr::filter(Age < 10)
coral <- coral[(coral$Calcite <= 1 | is.na(coral$Calcite)),]

#clean up a nice dataframe
coral.df <- coral %>% mutate(Temperature = Temp) %>%
  select(U238,pH,TAlk,Salinity,Temperature,OmegaA,TCO2) %>%
  data.frame()

coral.df$U238 <- coral.df$U238 * 0.421
nrow(coral.df)
```

```{r}
#This is a favia, so removing favia removes it
#coral[92,]
#coral[coral$Source==74,]
##obs 92 from site 74 is an extreme outlier
#coral.df <- coral.df[-92,]
```

```{r}
## three outliers
coral[c(114,647,661),]
```

## Data Distribution

Plot the distributions of the predictors. 
```{r distributions}
ggplot(gather(coral.df),aes(value))+
  geom_histogram(bins=10)+
  facet_wrap(~key, scales = 'free_x')
```

## Residaul Analysis

Residuals vs fitted looks perfectly normal - no signs of non-linearity or heteroskedasticity.

Normal Q-Q is somewhat concerning. According the shapiro test we reject the null hypothesis and conclude the data is non-normal.

Scale-location is maybe slightly non-linear, but not to an extent that is cause for concern. We can also formally test heteroskedasticity with a Breusch-Pagan test. The null hypothesis is homoskedasticity, and we fail to reject it.

There are no conerning observations when looking at cook's distance in our residuals vs leverage plot.

```{r residuals}
options(scipen=999)
final_model <- lm(U238 ~ pH + Temperature + OmegaA + TAlk, data=coral.df)
summary(final_model)
plot(final_model)
```

```{r additional}
#normal Q-Q additional analysis
par(mfrow=c(1,1))
d<-density(final_model[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
shapiro.test(final_model[['residuals']])

#scale-location additional analysis
bptest(final_model)
```
